Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-11-09 Manuscript/manuscript.tex
Small edits to Introduction

% Here I write pseudo-article statements that will make the main argument.
% Beautiful polished sentences will be formed only afte we agree on these basic things
\section{Introduction}
Ions interactions with cellular membranes play a key role in critical biological processes~\cite{REF_REVIEW}. Ions, especially cations, modify general properties of the membrane which ultimately modulate their embedded transmembrane proteins~\cite{REF}. Ions also have a more direct effect by modulating lipid--, protein--, and sugar--lipid interactions at the membrane surface. For example, Ca$^{2+}$  is crucial in neural signal propagation. It promotes membrane fusion by bridging lipids in the vesicles carrying neurotransmitters and the neuron synapsis membrane~\cite{REF}. Ca$^{2+}$ also participates in the T-cell receptor activation. It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for Lck protein~\cite{Shi2013}. A final example involving sugar-lipid interactions mediated by ions is that Ca$^{2+}$ can modulate the presentation of the sugars found in the PI(4,5)P$_2$ lipid which ultimately modulates phospholipase C delta 1 pleckstrin homology domain (PLC $\delta$1-H)~\cite{Bilkova2017Calcium}. Interestingly, this modulation is lost when using Mg$^{2+}$ instead of Ca$^{2+}$ illustrating the selectivity of these processes towards particular ions. Despite our increasing understanding of the role of ions in cell membrane-related processes the exact molecular details of the mechanisms behind such processes remain elusive.

Direct measurements of ion-membrane interactions in real biological systems are difficult. Simplified lipid bilayers are often used instead as entry-level models to shed light on the role of ions in complex biological membranes~\cite{scherer87,seelig90,cevc90}. For this reason, interactions of biologically relevant cations, especially Na$^+$ and Ca$^{2+}$, with zwitterionic phosphocholine (PC) bilayers have been widely studied in experiments~\cite{akutsu81, altenbach84, seelig90, cevc90, tocanne90, binder02, pabst07, uhrikova08} and classical MD simulations~\cite{bockmann03, bockmann04, Berkowitz12, melcrova16, javanainen17}. The details of ion binding are, however, not fully consistent in the literature. Non-invasive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering, and infrared spectroscopies mainly suggest that Na$^+$ ions exhibit negligible binding to PC lipid bilayers. On the contrary, Ca$^{2+}$ is observed to specifically bind to a couple of PC molecules using their phosphate groups~\cite{hauser76, hauser78, herbette84, akutsu81, altenbach84, binder02, pabst07, uhrikova08}. On the other hand, {\color{red} in general} \todo{Not CHARMM36 or amber14} atomistic resolution molecular dynamics (MD) simulation models predict a stronger binding for the cations~\cite{catte16}. For example, simulations report Na$^+$ accumulating in the lipid interface~\cite{bockmann03} or Ca$^{2+}$ binding up to 4 PC lipids simultaneously, including not only interactions with phosphates but also with carbonyl oxygens~\cite{bockmann04, melcrova16, javanainen17}. {\color{red} Also, several experiments have also been interpreted to support the predictions from earlier MD simulations~\cite{bockmann03, vacha09a}. \todo{remove? Why we need this?}}

Recent studies within the NMRlipids project \url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these apparent controversies. A direct comparison of ion binding affinities to PC bilayers between simulations and experiments was performed using the electrometer concept. Namely, the changes in NMR order parameters of the headgroup upon ions additions~\cite{seelig87} is directly compared to the MD simulations results. Analyzing massive amounts of data collected by the Open Collaboration method, it was concluded that the accuracy of the current state of the art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.

>>>>

% Here I write pseudo-article statements that will make the main argument.
% Beautiful polished sentences will be formed only afte we agree on these basic things
\section{Introduction}
Ion interactions with cellular membranes play a key role in critical biological processes~\cite{seelig90,cevc90}. Ions, especially multivalent cations, modify general properties of the membrane which ultimately modulate their embedded transmembrane proteins~\cite{cevc90,tocanne90, pabst07}. Ions also have a more direct effect by modulating lipid--, protein--, and sugar--lipid interactions at the membrane surface. For example, Ca$^{2+}$  is crucial in neural signal propagation. It promotes membrane fusion by bridging lipids in the vesicles carrying neurotransmitters and the neuron synapsis membrane~\cite{REF}. Ca$^{2+}$ also participates in the T-cell receptor activation. It induces the detachment of the positively charged cytosolic tails of the CD3 protein complex from the negatively charged intracellular membrane making it accessible for Lck protein~\cite{Shi2013}. A final example involving sugar-lipid interactions mediated by ions is that Ca$^{2+}$ can modulate the presentation of the sugars found in the PI(4,5)P$_2$ lipid which ultimately modulates phospholipase C delta 1 pleckstrin homology domain (PLC $\delta$1-H)~\cite{Bilkova2017Calcium}. Interestingly, this modulation is lost when using Mg$^{2+}$ instead of Ca$^{2+}$ illustrating the selectivity of these processes towards particular ions. Despite our increasing understanding of the role of ions in cell membrane-related processes the exact molecular details of the mechanisms behind such processes remain elusive.

Direct measurements of ion-membrane interactions in real biological systems are difficult. Simplified lipid bilayers are often used instead as entry-level models to shed light on the role of ions in complex biological membranes~\cite{scherer87,seelig90,cevc90}. For this reason, interactions of biologically relevant cations, especially Na$^+$ and Ca$^{2+}$, with zwitterionic phosphocholine (PC) bilayers have been widely studied in experiments~\cite{akutsu81, altenbach84, seelig90, cevc90, tocanne90, binder02, pabst07, uhrikova08} and classical MD simulations~\cite{bockmann03, bockmann04, Berkowitz12, melcrova16, javanainen17}. The details of ion binding are, however, not fully consistent in the literature. Non-invasive spectroscopic methods, like nuclear magnetic resonance (NMR), scattering, and infrared spectroscopies mainly suggest that Na$^+$ ions exhibit negligible binding to PC lipid bilayers. On the contrary, Ca$^{2+}$ is observed to specifically bind to a couple of PC molecules using their phosphate groups~\cite{hauser76, hauser78, herbette84, akutsu81, altenbach84, binder02, pabst07, uhrikova08}. On the other hand, in general atomistic resolution molecular dynamics (MD) simulation models predict a stronger binding for the cations~\cite{catte16}. For example, simulations report Na$^+$ accumulating in the lipid interface~\cite{bockmann03} or Ca$^{2+}$ binding up to 4 PC lipids simultaneously, including not only interactions with phosphates but also with carbonyl oxygens~\cite{bockmann04, melcrova16, javanainen17}.
%{\color{red} Also, several experiments have also been interpreted to support the predictions from earlier MD simulations~\cite{bockmann03, vacha09a}. \todo{remove? Why we need this?}}

Recent studies within the NMRlipids project (\url{nmrlipids.blogspot.fi})~\cite{catte16} made an attempt to resolve these apparent controversies. A direct comparison of ion binding affinities to PC bilayers between simulations and experiments was performed using the electrometer concept~\cite{seelig87}. Namely, the changes in NMR order parameters of the headgroup upon addition of ions is directly compared to the MD simulations results. Analyzing massive amounts of data collected by the Open Collaboration method, it was concluded that the accuracy of the current state of the art lipid models for MD simulations is not sufficient for a detailed interpretation of the interactions of cations with PC lipid bilayers~\cite{catte16}.
